---
title: "Assignment8"
output: ''
date: '2022-03-23'
---
#Loading Libraries
```{r}
library(here)
library(tidyverse)
library(devtools)
library(survey)
library(srvyr)
library(od)
library(scenRios)
library(sf)
library(ggplot2)
library(ggspatial)
library(lwgeom)
library(PNWColors)
```

```{r}
install_github("https://github.com/c-voulgaris/scenRios")
```
#Load trip generation data and skims
```{r}
zones <- here("Assignment7",
              "existing_zones.csv") %>%
  read_csv(show_col_types = FALSE)

proposed_zones <- here("Assignment7",
                       "proposed_zones.csv") %>%
  read_csv(show_col_types = FALSE)

skims <- here("existing",
              "data",
              "skims.csv") %>%
  read_csv(show_col_types = FALSE)
```

#Calculate average travel time by trip purpose
```{r}
temp <- tempfile()
download.file("https://nhts.ornl.gov/assets/2016/download/csv.zip", temp)

trips <- read_csv(unz(temp, "trippub.csv"), 
                      show_col_types = FALSE) %>%
  # 15380 is the GEOID for Buffalo - should be 19740 for Denver
  filter(HH_CBSA == "19740")

unlink(temp)
```
##Trip Purpose Variable Creation
```{r}
trips <- trips %>%
  mutate(home_based = case_when(WHYTO == "01" ~ TRUE,
                                WHYTO == "02" ~ TRUE,
                                WHYFROM == "01" ~ TRUE,
                                WHYFROM == "02" ~ TRUE,
                                TRUE ~ FALSE)) %>%
  mutate(work = ifelse(WHYTO == "03" | WHYFROM == "03", TRUE, FALSE)) %>%
  mutate(purpose = case_when(home_based & work ~ "HBW",
                            home_based ~ "HBO",
                            TRUE ~ "NHB"))
```
##Survey Object and Average Travel Time
```{r}
trips_svy <- trips %>%
  as_survey(weights = WTTRDFIN)

ttime_by_purpose <- trips_svy %>%
  group_by(purpose) %>%
  summarise(avg_time = survey_mean(TRVLCMIN))

ttime_by_purpose
```

#Calculating minumum travel time across all modes
```{r}
skims <- skims %>%
  mutate(min_time = pmin(transit_time, 
                         car_time,
                         bicycle_time,
                         walk_time,
                         na.rm = TRUE))
```

#Calculating Friction Factors
```{r}
skims <- skims %>%
  mutate(F_HBW = min_time^-0.503*exp(-0.078*min_time),
         F_HBO = min_time^-3.993*exp(-0.019*min_time),
         F_NHB = min_time^-3.345*exp(-0.003*min_time)) 
```

### EXISTING CONDITIONS ###

#HBW Travel Flows
```{r}
HBW_dist <- grvty_balancing(od_zones = zones,
                            friction = skims,
                            zone_id = "GEOID",
                            zone_o = "hbw_prod",
                            zone_d = "hbw_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_HBW",
                            tolerance = .01,
                            max_iter = 150)
```

```{r}
save(HBW_dist, file = here("Assignment8",
                            "HBW_dist"))
```

#NHB Travel Flows
```{r}
NHB_dist <- grvty_balancing(od_zones = zones,
                            friction = skims,
                            zone_id = "GEOID",
                            zone_o = "nhb_prod",
                            zone_d = "nhb_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_NHB",
                            tolerance = .01,
                            max_iter = 15000)
```

```{r}
save(NHB_dist, file = here("Assignment8",
                    "NHB_dist"))
```

#HBO Travel Flows
```{r}
HBO_dist <- grvty_balancing(od_zones = zones,
                            friction = skims,
                            zone_id = "GEOID",
                            zone_o = "hbo_prod",
                            zone_d = "hbo_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_HBO",
                            tolerance = .01,
                            max_iter = 150)
```

```{r}
save(HBO_dist, file = here("Assignment8",
                    "HBO_dist"))
```

#Convergence
##HBW Convergence
```{r}
tail(HBW_dist$convergence)
HBW_dist$convergence
```
##NHB Convergence
```{r}
tail(NHB_dist$convergence)
NHB_dist$convergence
```  
#HBO Convergence
```{r}
tail(HBO_dist$convergence)
HBO_dist$convergence
```

#Flows
##HBW Flows
```{r}
head(HBW_dist$flows)
table(HBW_dist$flows$flow > 0)
```
##NHB Flows
```{r}
head(NHB_dist$flows)
table(NHB_dist$flows$flow > 0)
```
##HBO Flows
```{r}
head(HBO_dist$flows)
table(HBO_dist$flows$flow > 0)
```

#Visualizing Convergence
```{r}
convergence_points_HBO <- HBO_dist$convergence %>%
  mutate(max_diff = max_o_diff + max_d_diff) %>%
  mutate(which_max = ifelse(max_o_diff > max_d_diff, 
                            "Productions",
                            "Attractions"))

hbo_convergence <- ggplot(convergence_points_HBO) +
  geom_line(aes(x = iteration, y = max_diff, lty = which_max)) +
  scale_y_continuous(name = "Maximum difference from target value",
                     trans = "log", 
                     breaks = breaks <- 10^seq(1,5, by=1),
                     labels = formatC(breaks, 
                                      big.mark = ",", 
                                      format = "f", 
                                      digits = 0)) +
  scale_x_continuous(name = "Iteration",
                     breaks = breaks <- seq(0, 40, by=10),
                     labels = formatC(breaks, 
                                      big.mark = ",", 
                                      format = "f", 
                                      digits = 0)) +
  scale_linetype(name = "") +
  theme_minimal()

ggsave("convergence_points_HBO.pdf", hbo_convergence)
```
```{r}
convergence_points_HBW <- HBW_dist$convergence %>%
  mutate(max_diff = max_o_diff + max_d_diff) %>%
  mutate(which_max = ifelse(max_o_diff > max_d_diff, 
                            "Productions",
                            "Attractions"))

hbw_convergence <- ggplot(convergence_points_HBW) +
  geom_line(aes(x = iteration, y = max_diff, lty = which_max)) +
  scale_y_continuous(name = "Maximum difference from target value",
                     trans = "log", 
                     breaks = breaks <- 10^seq(1,5, by=1),
                     labels = formatC(breaks, 
                                      big.mark = ",", 
                                      format = "f", 
                                      digits = 0)) +
  scale_x_continuous(name = "Iteration",
                     breaks = breaks <- seq(0, 20, by=5),
                     labels = formatC(breaks, 
                                      big.mark = ",", 
                                      format = "f", 
                                      digits = 0)) +
  scale_linetype(name = "") +
  theme_minimal()

ggsave("convergence_points_HBW.pdf", hbw_convergence)
```
```{r}
convergence_points_NHB <- NHB_dist$convergence %>%
  mutate(max_diff = max_o_diff + max_d_diff) %>%
  mutate(which_max = ifelse(max_o_diff > max_d_diff, 
                            "Productions",
                            "Attractions"))

nhb_convergence <- ggplot(convergence_points_NHB) +
  geom_line(aes(x = iteration, y = max_diff, lty = which_max)) +
  scale_y_continuous(name = "Maximum difference from target value",
                     trans = "log", 
                     breaks = breaks <- 10^seq(1,5, by=1),
                     labels = formatC(breaks, 
                                      big.mark = ",", 
                                      format = "f", 
                                      digits = 0)) +
  scale_x_continuous(name = "Iteration",
                     breaks = breaks <- seq(0, 25, by=5),
                     labels = formatC(breaks, 
                                      big.mark = ",", 
                                      format = "f", 
                                      digits = 0)) +
  scale_linetype(name = "") +
  theme_minimal()

ggsave("convergence_points_NHB.pdf", nhb_convergence)
```

#Mapping Desire Lines
##HBO Flows
```{r}
zone_boundaries <- here("zones",
              "boundaries.geojson") %>%
  st_read(quiet = TRUE)

desire_lines_HBO <- od_to_sf(HBO_dist$flows, zone_boundaries, silent = TRUE) %>%
  filter(flow > 0)

ggplot(desire_lines_HBO) +
  annotation_map_tile(type = "cartolight", zoom = 10, progress = "none") +
  geom_sf(aes(alpha = flow)) +
  theme_void()
```

```{r}
desire_lines_top_zones_HBO <- desire_lines_HBO %>%
  filter(d_id == c("08031001702","08031001701", "08001008100", "08005006856", "08005006858"))

fivezones_HBO <- ggplot(desire_lines_top_zones_HBO) +
  annotation_map_tile(type = "cartolight", zoom = 11, progress = "none") +
  geom_sf(aes(alpha = flow)) +
  theme_void()

ggsave("flows_hbo.pdf", fivezones_HBO)
```
##NHB Flows
```{r}
zone_boundaries <- here("zones",
              "boundaries.geojson") %>%
  st_read(quiet = TRUE)

desire_lines_NHB <- od_to_sf(NHB_dist$flows, zone_boundaries, silent = TRUE) %>%
  filter(flow > 0)

desire_lines_five_zones_NHB <- desire_lines_NHB %>%
  filter(d_id == c("08031001702","08031001701", "08005006856", "08001008100", "08031003800"))

fivezones_NHB <- ggplot(desire_lines_five_zones_NHB) +
  annotation_map_tile(type = "cartolight", zoom = 10, progress = "none") +
  geom_sf(aes(alpha = flow)) +
  theme_void()

ggsave("flows_nhb.pdf", fivezones_NHB)
```

##HBW Flows
```{r}
zone_boundaries <- here("zones",
              "boundaries.geojson") %>%
  st_read(quiet = TRUE)

desire_lines_HBW <- od_to_sf(HBW_dist$flows, zone_boundaries, silent = TRUE) %>%
  filter(flow > 0)

desire_lines_five_zones_HBW <- desire_lines_HBW %>%
  filter(d_id == c("08031001702","08031001701", "08005006856", "08005006712", "08001008100"))

fivezones_HBW <- ggplot(desire_lines_five_zones_HBW) +
  annotation_map_tile(type = "cartolight", zoom = 10, progress = "none") +
  geom_sf(aes(alpha = flow)) +
  theme_void()

ggsave("flows_hbw.pdf", fivezones_HBW)
```

#Comparing Calculated to Observed Times
```{r}
ttime_by_purpose
```

```{r}
HBO_flows <- HBO_dist$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         HBO_flow = flow)
HBW_flows <- HBW_dist$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         HBW_flow = flow)
NHB_flows <- NHB_dist$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         NHB_flow = flow)
```
```{r}
ex_skims_update <- skims %>%
  left_join(HBO_flows) %>%
  left_join(HBW_flows) %>%
  left_join(NHB_flows) %>%
  replace_na(list(HBO_flow = 0)) %>%
  replace_na(list(NHB_flow = 0)) %>%
  replace_na(list(HBW_flow = 0)) %>%
  mutate(total_time_HBO = HBO_flow * min_time,
         total_time_NHB = NHB_flow * min_time,
         total_time_HBW = HBW_flow * min_time)
```

```{r}

write.csv(ex_skims_update, here("existing",
                                "data",
                                "ex_skims_update.csv"))
```

```{r}
HBO_mean_time <- sum(ex_skims_update$total_time) / sum(ex_skims_update$HBO_flow)

HBO_mean_time
```

### ALTERNATIVE SECTION ###

##Alt HBW Travel Flows
```{r}
HBW_dist_alt <- grvty_balancing(od_zones = proposed_zones,
                            friction = ex_skims_update,
                            zone_id = "GEOID",
                            zone_o = "hbw_prod",
                            zone_d = "hbw_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_HBW",
                            tolerance = .01,
                            max_iter = 150)
```

```{r}
save(HBW_dist_alt, file = here("Assignment8",
                            "HBW_dist_alt"))
```

##Alt NHB Travel Flows
```{r}
NHB_dist_alt <- grvty_balancing(od_zones = proposed_zones,
                            friction = ex_skims_update,
                            zone_id = "GEOID",
                            zone_o = "nhb_prod",
                            zone_d = "nhb_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_NHB",
                            tolerance = .01,
                            max_iter = 150)
```

```{r}
save(NHB_dist, file = here("Assignment8",
                    "NHB_dist_alt"))
```

##Alt HBO Travel Flows
```{r}
HBO_dist_alt <- grvty_balancing(od_zones = proposed_zones,
                            friction = ex_skims_update,
                            zone_id = "GEOID",
                            zone_o = "hbo_prod",
                            zone_d = "hbo_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_HBO",
                            tolerance = .01,
                            max_iter = 150)
```

```{r}
save(HBO_dist, file = here("Assignment8",
                    "HBO_dist_alt"))
```

#Visualizing Proposed Convergence
```{r}
convergence_points_HBO_alt <- HBO_dist_alt$convergence %>%
  mutate(max_diff = max_o_diff + max_d_diff) %>%
  mutate(which_max = ifelse(max_o_diff > max_d_diff, 
                            "Productions",
                            "Attractions"))

hbo_convergence_alt <- ggplot(convergence_points_HBO_alt) +
  geom_line(aes(x = iteration, y = max_diff, lty = which_max)) +
  scale_y_continuous(name = "Maximum difference from target value",
                     trans = "log", 
                     breaks = breaks <- 10^seq(1,5, by=1),
                     labels = formatC(breaks, 
                                      big.mark = ",", 
                                      format = "f", 
                                      digits = 0)) +
  scale_x_continuous(name = "Iteration",
                     breaks = breaks <- seq(0, 40, by=10),
                     labels = formatC(breaks, 
                                      big.mark = ",", 
                                      format = "f", 
                                      digits = 0)) +
  scale_linetype(name = "") +
  theme_minimal()

ggsave("convergence_points_HBO_alt.pdf", hbo_convergence_alt)
```

```{r}
convergence_points_HBW_alt <- HBW_dist_alt$convergence %>%
  mutate(max_diff = max_o_diff + max_d_diff) %>%
  mutate(which_max = ifelse(max_o_diff > max_d_diff, 
                            "Productions",
                            "Attractions"))

hbw_convergence_alt <- ggplot(convergence_points_HBW_alt) +
  geom_line(aes(x = iteration, y = max_diff, lty = which_max)) +
  scale_y_continuous(name = "Maximum difference from target value",
                     trans = "log", 
                     breaks = breaks <- 10^seq(1,5, by=1),
                     labels = formatC(breaks, 
                                      big.mark = ",", 
                                      format = "f", 
                                      digits = 0)) +
  scale_x_continuous(name = "Iteration",
                     breaks = breaks <- seq(0, 20, by=5),
                     labels = formatC(breaks, 
                                      big.mark = ",", 
                                      format = "f", 
                                      digits = 0)) +
  scale_linetype(name = "") +
  theme_minimal()

ggsave("convergence_points_HBW_alt.pdf", hbw_convergence_alt)
```

```{r}
convergence_points_NHB_alt <- NHB_dist_alt$convergence %>%
  mutate(max_diff = max_o_diff + max_d_diff) %>%
  mutate(which_max = ifelse(max_o_diff > max_d_diff, 
                            "Productions",
                            "Attractions"))

nhb_convergence_alt <- ggplot(convergence_points_NHB_alt) +
  geom_line(aes(x = iteration, y = max_diff, lty = which_max)) +
  scale_y_continuous(name = "Maximum difference from target value",
                     trans = "log", 
                     breaks = breaks <- 10^seq(1,5, by=1),
                     labels = formatC(breaks, 
                                      big.mark = ",", 
                                      format = "f", 
                                      digits = 0)) +
  scale_x_continuous(name = "Iteration",
                     breaks = breaks <- seq(0, 25, by=5),
                     labels = formatC(breaks, 
                                      big.mark = ",", 
                                      format = "f", 
                                      digits = 0)) +
  scale_linetype(name = "") +
  theme_minimal()

ggsave("convergence_points_NHB_alt.pdf", nhb_convergence_alt)
```

#Mapping Desire Lines
##HBO Flows
```{r}
zone_boundaries <- here("zones",
              "boundaries.geojson") %>%
  st_read(quiet = TRUE)

desire_lines_HBO_alt <- od_to_sf(HBO_dist_alt$flows, zone_boundaries, silent = TRUE) %>%
  filter(flow > 0)

desire_lines_top_zones_HBO_alt <- desire_lines_HBO_alt %>%
  filter(d_id == c("08031002702","08031003800", "08031001702", "08031004301", "08005006856"))

fivezones_HBO_alt <- ggplot(desire_lines_top_zones_HBO_alt) +
  annotation_map_tile(type = "cartolight", zoom = 11, progress = "none") +
  geom_sf(aes(alpha = flow)) +
  theme_void()

ggsave("flows_hbo_alt.pdf", fivezones_HBO_alt)
```
##NHB Flows
```{r}
zone_boundaries <- here("zones",
              "boundaries.geojson") %>%
  st_read(quiet = TRUE)

desire_lines_NHB_alt <- od_to_sf(NHB_dist_alt$flows, zone_boundaries, silent = TRUE) %>%
  filter(flow > 0)

desire_lines_five_zones_NHB_alt <- desire_lines_NHB_alt %>%
  filter(d_id == c("08031002702","08031003800", "08031001702", "08005006856", "08031004301"))

fivezones_NHB_alt <- ggplot(desire_lines_five_zones_NHB_alt) +
  annotation_map_tile(type = "cartolight", zoom = 10, progress = "none") +
  geom_sf(aes(alpha = flow)) +
  theme_void()

ggsave("flows_nhb_alt.pdf", fivezones_NHB_alt)
```

##HBW Flows
```{r}
zone_boundaries <- here("zones",
              "boundaries.geojson") %>%
  st_read(quiet = TRUE)

desire_lines_HBW_alt <- od_to_sf(HBW_dist_alt$flows, zone_boundaries, silent = TRUE) %>%
  filter(flow > 0)

desire_lines_five_zones_HBW_alt <- desire_lines_HBW_alt %>%
  filter(d_id == c("08031002702","08031001702", "08005006856", "08031004301", "08031003800"))

fivezones_HBW_alt <- ggplot(desire_lines_five_zones_HBW_alt) +
  annotation_map_tile(type = "cartolight", zoom = 10, progress = "none") +
  geom_sf(aes(alpha = flow)) +
  theme_void()

ggsave("flows_hbw_alt.pdf", fivezones_HBW_alt)
```

```{r}
HBO_flows_alt <- HBO_dist_alt$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         HBO_flow = flow)
HBW_flows_alt <- HBW_dist_alt$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         HBW_flow = flow)
NHB_flows_alt <- NHB_dist_alt$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         NHB_flow = flow)
```
```{r}
alt_skims_update <- skims %>%
  left_join(HBO_flows_alt) %>%
  left_join(HBW_flows_alt) %>%
  left_join(NHB_flows_alt) %>%
  replace_na(list(HBO_flow = 0)) %>%
  replace_na(list(NHB_flow = 0)) %>%
  replace_na(list(HBW_flow = 0)) %>%
  mutate(total_time_HBO = HBO_flow * min_time,
         total_time_NHB = NHB_flow * min_time,
         total_time_HBW = HBW_flow * min_time)
```

```{r}

write.csv(alt_skims_update, here("alternative",
                                "data",
                                "alt_skims_update.csv"))
```
#Looking at difference between existing and proposed trip flows

```{r} 
#thanks for the inspiration San Jose!
#combining desire line
desire_lines_all <- desire_lines_HBO %>%
  rename(hbo_flow = flow) %>%
  as.data.frame() %>%
  full_join(desire_lines_HBW %>%
              rename(hbw_flow = flow) %>%
              as.data.frame(), 
            by = c("o_id", "d_id", "geometry")) %>%
  full_join(desire_lines_NHB %>%
              rename(nhb_flow = flow) %>%
              as.data.frame(),
            by = c("o_id", "d_id", "geometry")) %>%
  st_as_sf()
```
```{r}
desire_lines_all_alt <- desire_lines_HBO_alt %>%
  rename(hbo_flow_alt = flow) %>%
  as.data.frame() %>%
  full_join(desire_lines_HBW_alt %>%
              rename(hbw_flow_alt = flow) %>%
              as.data.frame(), 
            by = c("o_id", "d_id", "geometry")) %>%
  full_join(desire_lines_NHB_alt %>%
              rename(nhb_flow_alt = flow) %>%
              as.data.frame(),
            by = c("o_id", "d_id", "geometry")) %>%
  st_as_sf()
```
#Joining existing and proposed
```{r}
desire_lines_all_combined <- desire_lines_all %>%
  as.data.frame() %>%
  full_join(desire_lines_all_alt %>%
              as.data.frame(),
            by = c("o_id", "d_id", "geometry")) %>%
  mutate(hbo_flow_dif = (hbo_flow_alt - hbo_flow) / hbo_flow,
         hbw_flow_dif = (hbw_flow_alt - hbw_flow) / hbw_flow,
         nhb_flow_dif = (nhb_flow_alt - nhb_flow) / nhb_flow) %>%
  st_as_sf()
```

```{r}
trip_pal <- pnw_palette(name = "Bay", 
                           n = 6, 
                           type = "continuous")
```

```{r}
diff_NHB <- od_to_sf(desire_lines_all_combined, zone_boundaries, silent = TRUE) %>%
  filter(o_id == "08001009206") %>%
  filter(nhb_flow_dif != 0)

diff_NHB_map <- ggplot(diff_NHB) +
  annotation_map_tile(type = "cartolight", zoom = 10, progress = "none") +
  geom_sf(aes(color = nhb_flow_dif)) +
  scale_color_gradient2(low = "red", high = "blue", mid = "grey", midpoint = 0, trans = "pseudo_log") +
  theme_void()

diff_NHB_map

ggsave("diff_NHB.pdf", diff_NHB_map)
```

```{r}
diff_HBW <- od_to_sf(desire_lines_all_combined, zone_boundaries, silent = TRUE) %>%
  filter(o_id == "08059060500") %>%
  filter(hbw_flow_dif != 0)

diff_HBW_map <- ggplot(diff_HBW) +
  annotation_map_tile(type = "cartolight", zoom = 10, progress = "none") +
  geom_sf(aes(color = hbw_flow_dif)) +
  scale_color_gradient2(low = "red", high = "blue", mid = "grey", midpoint = 0, trans = "pseudo_log") +
  theme_void()

diff_HBW_map

ggsave("diff_HBW.pdf", diff_HBW_map)
```
```{r}
diff_HBO <- od_to_sf(desire_lines_all_combined, zone_boundaries, silent = TRUE) %>%
  filter(o_id == "08001009204") %>%
  filter(hbo_flow_dif != 0)

diff_HBO_map <- ggplot(diff_HBO) +
  annotation_map_tile(type = "cartolight", zoom = 10, progress = "none") +
  geom_sf(aes(color = hbo_flow_dif)) +
  scale_color_gradient2(low = "red", high = "blue", mid = "grey", midpoint = 0, trans = "pseudo_log") +
  theme_void()

ggsave("diff_HBO.pdf", diff_HBO_map)
```