---
title: "Assignment_6"
author: "Miguel Perez Luna"
date: "3/2/2022"
output: html_document
---

```{r}
library(here)
library(tidyverse)
library(sf)
library(modelr)
library(jtools)
```

# Loading in data

We are loading our accessibility data for our existing conditions and our zone data. 

```{r}
access <- here("existing/data/access_compare_existing.csv") %>%
  read_csv()

zones <- here("existing/data/zone_data.csv") %>%
  read_csv()
```

# Selecting Variables

```{r}
model_data <- left_join(zones, access) %>%
  mutate(pct_smol_hh = (oneperson_hh + twoperson_hh) / housing_units,
         pct_lo_inc = hh_income_low / hh_total,
         pct_hi_inc = hh_income_high / hh_total,
         pct_no_veh = nocar / housing_units) %>%
  select(GEOID, pct_smol_hh, pct_lo_inc, pct_hi_inc, ratio, pct_no_veh)
```

```{r}
model_data[is.na(model_data) | model_data == "Inf"] <- NA

model_data <- na.omit(model_data) 
```

# Visualizing model variables

## No-Car Households vs. Low-income Households

```{r}
ggplot(model_data, aes(x = pct_lo_inc, y = pct_no_veh)) +
  geom_point(size = 1, color = "blue") +
  geom_smooth(size = 0.5, lty = "dashed", color = "blue", method = "lm")
```

## Distribution of Low-Income Households

```{r}
pct_lo_inc_dist <- ggplot(model_data) +
  geom_histogram(aes(x = pct_lo_inc))

pct_lo_inc_dist
```

## No-Car Households vs. High-Income Households

```{r}
ggplot(model_data, aes(x = pct_hi_inc, y = pct_no_veh)) +
  geom_point(size = 1, color = "red") +
  geom_smooth(size = 0.5, lty = "dashed", color = "red", method = "lm")
```

## No-Car Households vs. 1 to 2 person Households

```{r}
ggplot(model_data, aes(x = pct_smol_hh, y = pct_no_veh)) +
  geom_point(size = 1, color = "green") +
  geom_smooth(size = 0.5, lty = "dashed", color = "green", method = "lm")
```

## No-Car Households vs. Transit to Car Accessibility Ratio

```{r}
ggplot(model_data, aes(x = ratio, y = pct_no_veh)) +
  geom_point(size = 1, color = "black") +
  xlim(0,1) +
  geom_smooth(size = 0.5, lty = "dashed", color = "black", method = "lm")
```

# Estimating the regression model

```{r}
model1 <- lm(pct_no_veh ~ pct_smol_hh + pct_lo_inc + pct_hi_inc + ratio, data = model_data)

export_summs(model1)
```

## We are making a more parsimonious model.

```{r}
model2 <- lm(pct_no_veh ~ pct_smol_hh + pct_lo_inc, data = model_data)

export_summs(model2)
```

# Applying our model

Here we are reading in our alternative conditions zone data.

```{r}
alt_zone_data <- here("alternative",
                    "data",
                    "final_tracts_emp.csv") %>%
                read_csv() %>%
  mutate(new_total_hh = new_hh_income_low + new_hh_income_mid + new_hh_income_high)
```

Conundrum: We are not sure if we should be using our existing conditions accessibility ("access") in the left_join() function below. We are also not sure if we should be calling existing or proposed variables into the mutate() function. 


```{r}
alt_access <- here("alternative",
               "data",
               "access_compare_alt.csv") %>%
  read_csv()

alt_zones <- here("alternative",
                    "data",
                    "final_tracts_emp.csv") %>%
                read_csv() %>%
  mutate(new_total_hh = new_hh_income_low + new_hh_income_mid + new_hh_income_high) %>%
  left_join(alt_access) %>%
  mutate(pct_smol_hh = (new_oneperson_hh + new_twoperson_hh) / new_total_housing_units,
         pct_lo_inc = new_hh_income_low / new_total_hh,
         pct_hi_inc = new_hh_income_high / new_total_hh,
         pct_no_veh = new_nocar / new_total_housing_units) %>%
  add_predictions(model = model2) %>%
  mutate(new_pct_no_veh = pred) %>%
  mutate(new_nocar = round(new_total_housing_units * new_pct_no_veh)) 
```

