---
title: "assignment5_graphics"
output: html_document
---
## Load Libraries

```{r}
library(ggplot2)
library(sf)
library(tidyverse)
library(tidycensus)
library(sp)
library(units)
library(here)
library(png)
library(ggthemes)
library(RColorBrewer)
library(PNWColors)
```

## Reading in our .csv files

```{r}
car_access_existing <- read.csv(here("existing",
                                      "data",
                                      "car_access_existing.csv"))


walk_access_existing <- read.csv(here("existing",
                                      "data",
                                      "walk_access_existing.csv"))

bike_access_existing <- read.csv(here("existing",
                                      "data",
                                      "bike_access_existing.csv"))

transit_access_existing <- read.csv(here("existing",
                                      "data",
                                      "transit_access_existing.csv"))

car_access_alt <- read.csv(here("alternative",
                                      "data",
                                      "car_access_alt.csv"))

walk_access_alt <- read.csv(here("alternative",
                                      "data",
                                      "walk_access_alt.csv"))

bike_access_alt <- read.csv(here("alternative",
                                      "data",
                                      "bike_access_alt.csv"))

transit_access_alt <- read.csv(here("alternative",
                                      "data",
                                      "transit_access_alt.csv"))

access_compare_existing <- read.csv(here("existing",
                                      "data",
                                      "access_compare_existing.csv"))

access_compare_alt <- read.csv(here("alternative",
                                      "data",
                                      "access_compare_alt.csv"))
```

## Combining all modes of access files
```{r} 
# Looked to Austin group for guidance

all_access_existing <- left_join(car_access_existing, bike_access_existing) %>%
  left_join(walk_access_existing) %>%
  left_join(transit_access_existing) %>%
  left_join(access_compare_existing, by = c("from_id" = "GEOID"))

all_access_alt <- left_join(car_access_alt, bike_access_alt) %>%
  left_join(walk_access_alt) %>%
  left_join(transit_access_alt) %>%
  left_join(access_compare_alt, by = c("from_id" = "GEOID"))
```

## Joining combined modes with spatial data
```{r}
zones <- st_read(here("zones", "boundaries1.geojson")) %>%
  mutate(GEOID =
           as.numeric(GEOID))

all_access_existing <- left_join(zones, 
                                 all_access_existing, 
                                 by = c("GEOID" = "GEOID"))
  
all_access_alt <- left_join(zones,
                            all_access_alt,
                            by = c("GEOID" = "GEOID"))
```

```{r}
car_pal <- pnw_palette(name = "Bay", 
                           n = 6, 
                           type = "continuous")
```
## Making Maps

### Driving Under existing conditions

```{r}

car_access_alt_map <- ggplot(all_access_alt) +
  geom_sf(aes(fill = car_access),
          color = NA) +
  scale_fill_gradientn(colours = car_pal,
                       name = "Job Accessibility \nby Driving (Proposed)",
                       breaks = size_brks <- seq(0, 1000000, by = 200000),
                       labels = formatC(size_brks, format = "d", big.mark = ",")) +
  theme_void()

car_access_alt_map
```
### Walking under existing conditions

### Biking under existing conditions

### Transit under existing conditions

### Driving Under proposed conditions

### Walking under proposed conditions

### Biking under proposed conditions

### Transit under proposed conditions