---
title: "CalculatingAccessibility"
author: "Miguel Perez Luna"
date: "2/25/2022"
output: html_document
---
```{r}
options(java.parameters = '-Xmx4G')

library(r5r)
library(here)
library(tidyverse)
library(sf)
library(lubridate)
```

# Setting up Centroid Information for Existing Conditions

Here we are attaching information about the number of jobs at each destination to our set of centroids. 

```{r}

jobs <- here("existing",
             "data",
             "zone_data.csv") %>%
    read_csv() %>%
    select(GEOID, total_emp) %>%
    mutate(GEOID = as.character(GEOID)) %>%
    rename(id = GEOID)

centroids_jobs <- here("zones",
                  "centroids.geojson") %>%
    st_read() %>%
    left_join(jobs) 

```

# Setting up r5r

```{r}
existing_core <- here("existing",
                      "networks") %>%
  setup_r5(verbose = FALSE)
```

# Calculating accessibility for non-transit modes under existing conditions

## Calculating accessibility for cars

```{r}
car_access_existing <- accessibility(existing_core,
                                     origins = centroids_jobs,
                                     destinations = centroids_jobs,
                                     opportunities_colname = 'total_emp',
                                     mode = "CAR",
                                     decay_function = "logistic",
                                     cutoffs = 15,
                                     decay_value = 2, 
                                     verbose = FALSE)
```

```{r}
write_csv(car_access_existing, here("existing/data/car_access_existing.csv"))
```

## Calculating accessibility for pedestrians

```{r}
walk_access_existing <- accessibility(existing_core,
                                     origins = centroids_jobs,
                                     destinations = centroids_jobs,
                                     opportunities_colname = 'total_emp',
                                     mode = "WALK",
                                     decay_function = "logistic",
                                     cutoffs = 15,
                                     decay_value = 2, 
                                     verbose = FALSE)
```

```{r}
write_csv(walk_access_existing, here("existing/data/walk_access_existing.csv"))
```

## Calculating accessibility for bicyclists

```{r}
bike_access_existing <- accessibility(existing_core,
                                     origins = centroids_jobs,
                                     destinations = centroids_jobs,
                                     opportunities_colname = 'total_emp',
                                     mode = "BICYCLE",
                                     decay_function = "logistic",
                                     cutoffs = 15,
                                     decay_value = 2, 
                                     verbose = FALSE)
```

```{r}
write_csv(bike_access_existing, here("existing/data/bike_access_existing.csv"))
```

## Calculating accessibility for transit

```{r}
transit_access_existing <- accessibility(existing_core,
                                     origins = centroids_jobs,
                                     destinations = centroids_jobs,
                                     opportunities_colname = 'total_emp',
                                     mode = "TRANSIT",
                                     decay_function = "logistic",
                                     cutoffs = 15,
                                     decay_value = 2, 
                                     verbose = FALSE,
                                     departure_datetime = 
                                       ymd_hm("2022-02-22 17:00"),
                                     time_window = 120)
```

```{r}
write_csv(transit_access_existing, here("existing/data/transit_access_existing.csv"))
```

# Setting up Centroid Information for Proposed Conditions

# Recalculating New Total Employment

```{r}
final_tracts_emp <- final_tracts_sel %>%
mutate(new_total_emp = 
         new_basic +
         new_retail + 
         new_service)

write_csv(final_tracts_emp, 
          here("alternative",
               "data",
               "final_tracts_emp.csv"))
```


```{r}

jobs_proposed <- here("alternative",
             "data",
             "final_tracts_emp.csv") %>%
    read_csv() %>%
    select(GEOID, new_total_emp) %>%
    mutate(GEOID = as.character(GEOID)) %>%
    rename(id = GEOID)

centroids_jobs_alt <- here("zones",
                  "centroids.geojson") %>%
    st_read() %>%
    left_join(jobs_proposed) 

write_csv(jobs_proposed, here("alternative/data/jobs_proposed.csv"))
```

# Setting up r5r

```{r}
existing_core <- here("existing",
                      "networks") %>%
  setup_r5(verbose = FALSE)
```

# Calculating accessibility for non-transit modes under proposed conditions

## Calculating accessibility for cars

```{r}
car_access_alt <- accessibility(existing_core,
                                     origins = centroids_jobs_alt,
                                     destinations = centroids_jobs_alt,
                                     opportunities_colname = 'new_total_emp',
                                     mode = "CAR",
                                     decay_function = "logistic",
                                     cutoffs = 15,
                                     decay_value = 2, 
                                     verbose = FALSE)
```

```{r}
write_csv(car_access_alt, here("alternative/data/car_access_alt.csv"))
```

## Calculating accessibility for pedestrians

```{r}
walk_access_alt <- accessibility(existing_core,
                                     origins = centroids_jobs_alt,
                                     destinations = centroids_jobs_alt,
                                     opportunities_colname = 'new_total_emp',
                                     mode = "WALK",
                                     decay_function = "logistic",
                                     cutoffs = 15,
                                     decay_value = 2, 
                                     verbose = FALSE)
```

```{r}
write_csv(walk_access_alt, here("alternative/data/walk_access_alt.csv"))
```

## Calculating accessibility for bicyclists

```{r}
bike_access_alt <- accessibility(existing_core,
                                     origins = centroids_jobs_alt,
                                     destinations = centroids_jobs_alt,
                                     opportunities_colname = 'new_total_emp',
                                     mode = "BICYCLE",
                                     decay_function = "logistic",
                                     cutoffs = 15,
                                     decay_value = 2, 
                                     verbose = FALSE)
```

```{r}
write_csv(bike_access_alt, here("alternative/data/bike_access_alt.csv"))
```

## Calculating accessibility for transit

```{r}
transit_access_alt <- accessibility(existing_core,
                                     origins = centroids_jobs_alt,
                                     destinations = centroids_jobs_alt,
                                     opportunities_colname = 'new_total_emp',
                                     mode = "TRANSIT",
                                     decay_function = "logistic",
                                     cutoffs = 15,
                                     decay_value = 2, 
                                     verbose = FALSE,
                                     departure_datetime = 
                                       ymd_hm("2022-02-22 17:00"),
                                     time_window = 120)
```

```{r}
write_csv(transit_access_alt, here("alternative/data/transit_access_alt.csv"))
```

```{r}
stop_r5()
```

# Comparing auto and transit accessibility in existing conditions
```{r}
car_access_existing <- read_csv(here("existing/data/car_access_existing.csv"))
transit_access_existing <- read_csv(here("existing/data/transit_access_existing.csv"))
```

```{r}
car_access_existing_comp <- car_access_existing %>%
  select(from_id, accessibility) %>%
  rename(GEOID = from_id,
         car_access = accessibility)

transit_access_existing_comp <- transit_access_existing %>%
  select(from_id, accessibility) %>%
  rename(GEOID = from_id,
         transit_access = accessibility)

access_compare_existing <- left_join(car_access_existing_comp, transit_access_existing_comp) %>%
  mutate(ratio = transit_access / car_access)
```
```{r}
write_csv(access_compare_existing, here("existing/data/access_compare_existing.csv"))
```

# Comparing auto and transit accessibility in proposed conditions

First we are reading the saved csv files for Car and Transit accessibility

```{r}
car_access_alt <- read_csv(here("alternative/data/car_access_alt.csv"))
transit_access_alt <- read_csv(here("alternative/data/transit_access_alt.csv"))
```

Then we can calculate the quality of each by calculating a ratio. This is the ratio of transit accessibility to car accessibility. A ratio of 1 or higher means that there are that many more jobs accessible by transit than by car. So a ratio of 2 would mean that a if 50 jobs are accessible by cars, then 100 are accessible by transit. A ratio less than 1 means that there are that many fewer jobs accessible by transit than by car. A ratio of 0.25 would mean that if 100 jobs are accessible by car, then only 25 jobs are accessible by transit. 


```{r}
car_access_alt_comp <- car_access_alt %>%
  select(from_id, accessibility) %>%
  rename(GEOID = from_id,
         car_access = accessibility)

transit_access_alt_comp <- transit_access_alt %>%
  select(from_id, accessibility) %>%
  rename(GEOID = from_id,
         transit_access = accessibility)

access_compare_alt <- left_join(car_access_alt_comp, transit_access_alt_comp) %>%
  mutate(ratio = transit_access / car_access)
```

```{r}
write_csv(access_compare_alt, here("alternative/data/access_compare_alt.csv"))
```