---
title: "Assignment_7"
author: "Arnav Murulidhar"
date: "3/9/2022"
output: html_document
---

```{r}
library(here)
library(tidyverse)
library(survey)
library(srvyr)
library(naniar)
library(jtools)
library(knitr)
library(sf)
library(ggthemes)
library(RColorBrewer)
library(PNWColors)
```
#Loading NHTS data

```{r}
hhs <- here("large_files/csv/hhpub.csv") %>%
  read_csv()

trips <- here("large_files/csv/trippub.csv") %>%
  read_csv()
```

#Collecting relevant household level data 

```{r}
denver_hhs <- hhs %>%
  filter(HH_CBSA == "19740") %>%
  mutate(inc_levels_ = case_when(HHFAMINC == "01" ~ "low",
                               HHFAMINC == "02" ~ "low",
                               HHFAMINC == "03" ~ "low",
                               HHFAMINC == "04" ~ "mid",
                               HHFAMINC == "05" ~ "mid",
                               HHFAMINC == "06" ~ "mid",
                               HHFAMINC == "07" ~ "mid",
                               HHFAMINC == "08" ~ "high",
                               HHFAMINC == "09" ~ "high",
                               HHFAMINC == "10" ~ "high",
                               HHFAMINC == "11" ~ "high",
                               TRUE ~ "NA")) %>%
  mutate(size_ = case_when(HHSIZE == 1 ~ "one",
                              HHSIZE == 2 ~ "two",
                              TRUE ~ "NA")) %>%
  mutate(zero_veh_ = (HHVEHCNT == 0)) %>%
  replace_with_na(list(inc_levels_ = "NA")) %>%
  replace_with_na(list(size_ = "NA")) %>%
  select(HOUSEID, zero_veh_, size_, inc_levels_, WTHHFIN) 
```

#Collecting relevant trip level data

```{r}
trips_by_purpose <- trips %>% 
  filter(HH_CBSA == "19740") %>%
  select(HOUSEID, WHYFROM, WHYTO) %>%
  mutate(home_based = case_when(WHYTO == "01" ~ TRUE,
                                WHYTO == "02" ~ TRUE,
                                WHYFROM == "01" ~ TRUE,
                                WHYFROM == "02" ~ TRUE,
                                TRUE ~ FALSE)) %>%
  mutate(work = ifelse(WHYTO == "03" | WHYFROM == "03", TRUE, FALSE)) %>%
  mutate(purpose = case_when(home_based & work ~ "HBW",
                            home_based ~ "HBO",
                            TRUE ~ "NHB")) %>%
  group_by(HOUSEID, purpose) %>%
  summarize(n = n()) %>%
  pivot_wider(names_from = "purpose", values_from = "n") 
```
#Joining trip to household data
```{r}
hh_trips <- left_join(denver_hhs, trips_by_purpose) %>%
  replace_na(list(HBW = 0,
                  HBO = 0,
                  NHB = 0))
```
#Creating a survey object
```{r}
svy_trips <- hh_trips %>%
  as_survey(weights = WTHHFIN)
```

#Estimate a household-level regression model for HBO trips
```{r}
HBO_model1 <- svyglm(HBO ~ zero_veh_ + size_ + inc_levels_, svy_trips)

export_summs(HBO_model1, 
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = "Full model")
```
#Estimating with only household size for HBO trips
```{r}
HBO_model2 <- svyglm(HBO ~ size_, svy_trips)

export_summs(HBO_model1, HBO_model2,
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = c("Full model", "Reduced model"))
```

#Estimate a household-level regression model for HBW trips
```{r}
HBW_model1 <- svyglm(HBW ~ zero_veh_ + size_ + inc_levels_, svy_trips)

export_summs(HBW_model1, 
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = "Full model")
```
#Estimating with only household size for HBW trips
```{r}
HBW_model2 <- svyglm(HBW ~ inc_levels_, svy_trips)

export_summs(HBW_model1, HBW_model2,
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = c("Full model", "Reduced model"))
```
#Estimate a household-level regression model for NHB trips
```{r}
NHB_model1 <- svyglm(NHB ~ zero_veh_ + size_ + inc_levels_, svy_trips)

export_summs(NHB_model1, 
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = "Full model")
```
#Estimating with only household size for NHB trips
```{r}
NHB_model2 <- svyglm(NHB ~ inc_levels_, svy_trips)

export_summs(NHB_model1, NHB_model2,
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = c("Full model", "Reduced model"))
```
#Applying trip production model to zonal data
```{r}
existing_zones <- here("existing",
                       "data",
                       "zone_data.csv") %>%
  read_csv() %>%
  mutate(hbo_prod = hh_total * HBO_model2$coefficients["(Intercept)"] +
                    twoperson_hh * HBO_model2$coefficients["size_two"],
         hbw_prod = hh_total * HBW_model1$coefficients["(Intercept)"] +
                    hh_income_low * HBW_model1$coefficients["inc_levels_low"],
         nhb_prod = hh_total * NHB_model1$coefficients["(Intercept)"] +
                    hh_income_low * NHB_model1$coefficients["inc_levels_low"])
```
#Applying trip attractions
```{r}
existing_zones <- existing_zones %>%
  replace_na(list(basic_emp = 0,
                  retail_emp = 0, 
                  service_emp = 0)) %>%
  mutate(hbo_attr = 0.7 * hh_total +
                    0.7 * basic_emp +
                    8.4 * retail_emp +
                    3.5 * service_emp) %>%
  mutate(hbw_attr = 1.2 * (basic_emp + 
                             retail_emp + 
                             service_emp)) %>%
  mutate(nhb_attr = 0.6 * hh_total +
                    0.5 * basic_emp +
                    4.7 * retail_emp +
                    1.4 * service_emp)
```

#Comparing productions and attractions
```{r}
trip_end_summary <- tibble(Purpose = c("HBO"),
                           Productions = c(sum(existing_zones$hbo_prod)),
                           Attractions = c(sum(existing_zones$hbo_attr)),
                           Difference = c(sum(existing_zones$hbo_attr) - 
                                            sum(existing_zones$hbo_prod)))

kable(trip_end_summary, format.args = list(big.mark = ","))
```

```{r}
trip_end_summary <- tibble(Purpose = c("HBW"),
                           Productions = c(sum(existing_zones$hbw_prod)),
                           Attractions = c(sum(existing_zones$hbw_attr)),
                           Difference = c(sum(existing_zones$hbw_attr) - 
                                            sum(existing_zones$hbw_prod)))
kable(trip_end_summary, format.args = list(big.mark = ","))
```

```{r}
trip_end_summary <- tibble(Purpose = c("NHB"),
                           Productions = c(sum(existing_zones$nhb_prod)),
                           Attractions = c(sum(existing_zones$nhb_attr)),
                           Difference = c(sum(existing_zones$nhb_attr) - 
                                            sum(existing_zones$nhb_prod)))
kable(trip_end_summary, format.args = list(big.mark = ","))
```

#Balancing productions and attractions
```{r}
existing_zones <- existing_zones %>%
  mutate(hbo_attr_bal = hbo_attr * 
           sum(existing_zones$hbo_prod) / sum(existing_zones$hbo_attr)) %>%
  mutate(hbw_attr_bal = hbw_attr * 
           sum(existing_zones$hbw_prod) / sum(existing_zones$hbw_attr)) %>%
  mutate(nhb_attr_bal = nhb_attr * 
           sum(existing_zones$nhb_prod) / sum(existing_zones$nhb_attr))

write.csv(existing_zones, here("Assignment7/existing_zones.csv"))
```

#Comparing productions and attractions
```{r}
trip_end_summary <- tibble(Purpose = c("HBO"),
                           Productions = c(sum(existing_zones$hbo_prod)),
                           Attractions = c(sum(existing_zones$hbo_attr_bal)),
                           Difference = c(sum(existing_zones$hbo_attr_bal) - 
                                            sum(existing_zones$hbo_prod)))

kable(trip_end_summary, format.args = list(big.mark = ","))

trip_end_summary <- tibble(Purpose = c("HBW"),
                           Productions = c(sum(existing_zones$hbw_prod)),
                           Attractions = c(sum(existing_zones$hbw_attr_bal)),
                           Difference = c(sum(existing_zones$hbw_attr_bal) - 
                                            sum(existing_zones$hbw_prod)))

kable(trip_end_summary, format.args = list(big.mark = ","))

trip_end_summary <- tibble(Purpose = c("NHB"),
                           Productions = c(sum(existing_zones$nhb_prod)),
                           Attractions = c(sum(existing_zones$nhb_attr_bal)),
                           Difference = c(sum(existing_zones$nhb_attr_bal) - 
                                            sum(existing_zones$nhb_prod)))

kable(trip_end_summary, format.args = list(big.mark = ","))
```
#Creating a choropleth map
```{r}
existing_zones_select <- existing_zones %>% 
  select("GEOID", "hbo_prod", "hbw_prod", "nhb_prod", "hbo_attr_bal")

zones_pred <- here("zones",
                   "boundaries1.geojson") %>%
  st_read()


#left join zones and alt_zones

trip_geom <- left_join(zones_pred, existing_zones_select)
```

```{r}
trip_pal <- pnw_palette(name = "Bay", 
                           n = 6, 
                           type = "continuous")
```

```{r}
hbo_map <- ggplot(trip_geom) +
  geom_sf(aes(fill = hbo_prod), color=NA) +
   scale_fill_gradientn(colours = trip_pal,
                       name = "Tract home based other trips",
                       breaks = size_brks <- seq(0, 19000, by = 5000),
                       limits =  c(0,19000)) +
  theme_void()

hbo_map

ggsave("hbo_map.pdf" ,
         hbo_map)
```
```{r}
hbw_map <- ggplot(trip_geom) +
  geom_sf(aes(fill = hbw_prod), color=NA) +
   scale_fill_gradientn(colours = trip_pal,
                       name = "Tract home based work trips",
                       breaks = size_brks <- seq(0, 8000, by = 2000),
                       limits =  c(0,8000)) +
  theme_void()

hbw_map

ggsave("hbw_map.pdf" ,
         hbw_map)
```

```{r}
nhb_map <- ggplot(trip_geom) +
  geom_sf(aes(fill = nhb_prod), color=NA) +
   scale_fill_gradientn(colours = trip_pal,
                       name = "Tract non-home based trips",
                       breaks = size_brks <- seq(0, 12000, by = 3000),
                       limits =  c(0,12000)) +
  theme_void()

nhb_map

ggsave("nhb_map.pdf" ,
         nhb_map)
```

```{r}
hbo_attr_map <- ggplot(trip_geom) +
  geom_sf(aes(fill = hbo_attr_bal), color=NA) +
   scale_fill_gradientn(colours = trip_pal,
                       name = "Home based other trip attractions",
                       breaks = size_brks <- seq(0, 12000, by = 3000),
                       limits =  c(0,12000)) +
  theme_void()

hbo_attr_map

ggsave("hbo_attr_map.pdf" ,
         hbo_attr_map)
```

