---
title: "Comparing Existing and Proposed Conditions with Maps and Tables"
output: html_document
---
## Load Libraries

```{r}
library(ggplot2)
library(sf)
library(tidyverse)
library(tidycensus)
library(sp)
library(units)
library(here)
library(png)
library(ggthemes)
```

### Bringing in our Combined Census Tract Data

Here, we opened the csv file that we completed at the end of our second markdown file, TractZoneOverlay.Rmd. We also changed the values in the Proposed conditions variables to be numeric values.

```{r}
combinedtracts <- here("alternative",
                      "data",
                      "combined_cities",
                      "combined_tracts1.csv") %>%
  
  st_read() %>%
  mutate(new_basic = 
           as.numeric(new_basic),
         new_retail = 
           as.numeric(new_retail),
         new_service = 
           as.numeric(new_service),
         new_oneperson_hh = 
           as.numeric(new_oneperson_hh),
         new_twoperson_hh = 
           as.numeric(new_twoperson_hh),
         new_total_tract_population = 
           as.numeric(new_total_tract_population),
         new_total_housing_units = 
           as.numeric(new_total_housing_units),
         new_hh_income_low = 
           as.numeric(new_hh_income_low),
         new_hh_income_mid = 
           as.numeric(new_hh_income_mid),
         new_hh_income_high = 
           as.numeric(new_hh_income_high),
         new_nocar = 
           as.numeric(new_nocar))
         
  
```

### Selecting Variables to then Combine With Existing Conditions Data

To ensure that our new proposed condition variables would match with existing conditions data by tract, we created a new dataframe that contained the GEOID of our new proposed conditions and only the final counts of our new variables.

```{r}
final_tracts_sel <- combinedtracts %>%
  select(GEOID, 
         new_oneperson_hh,
         new_twoperson_hh,
         new_total_tract_population,
         new_total_housing_units,
         new_hh_income_low,
         new_hh_income_mid,
         new_hh_income_high,
         new_nocar,
         new_basic,
         new_retail,
         new_service)
```

### Combining Existing Conditions Tract Data with Proposed Conditions Tract Data

Last, we combined the Existing Conditions Tract Data from our first markdown file, ExistingConditions.Rmd, with the Proposed Conditions Data from our second markdown file, TractZoneOverlay.Rmd. They joined by GEOID.


```{r}
final_tracts <- left_join(zones, final_tracts_sel)
```

##Visualizing Existing and Proposed Conditions

Finally, we displayed comparison maps for each variable which compared the existing conditions of our tracts to the proposed conditions of our tracts. We did this using ggplot. We also saved each map as a pdf file for later addition to our report. 

###Population

####Existing Population

```{r}
existing_pop_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = population),color=NA)

existing_pop_map + 
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Tract Population",
                       breaks = size_brks <- seq(0, 40000, by = 10000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "People"))

ggsave("population_tract.pdf", 
  existing_pop_map + 
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Tract Population",
                       breaks = size_brks <- seq(0, 40000, by = 10000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "People")))
```
####Proposed Population

```{r}
proposed_pop_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_total_tract_population),color=NA)

proposed_pop_map + 
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Tract Population",
                       breaks = size_brks <- seq(0, 40000, by = 10000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "People"))

ggsave("proposed_population.pdf", 
  proposed_pop_map + 
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Tract Population",
                       breaks = size_brks <- seq(0, 40000, by = 10000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "People")))
```
###Households

####Existing Households
```{r}
existing_hh_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = housing_units),color=NA)

existing_hh_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Tract Housing Units",
                       breaks = size_brks <- seq(0, 7000, by = 1000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Units"))

ggsave("existing_hh.pdf", 
  existing_hh_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Tract Housing Units",
                       breaks = size_brks <- seq(0, 7000, by = 1000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Units")))
```

####Proposed Households

```{r}
proposed_hh_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_total_housing_units),color=NA)

proposed_hh_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Tract Housing Units",
                       breaks = size_brks <- seq(0, 15000, by = 2500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Units"))

ggsave("proposed_hh.pdf", 
  proposed_hh_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Tract Housing Units",
                       breaks = size_brks <- seq(0, 15000, by = 2500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Units")))
```

###Household Incomes

####Existing Low

```{r}
existing_lowincome_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = hh_income_low),color=NA)

existing_lowincome_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Low Income Households",
                       breaks = size_brks <- seq(0, 1200, by = 200),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("existing_lowincome.pdf", 
  existing_lowincome_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Low Income Households",
                       breaks = size_brks <- seq(0, 1200, by = 200),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

####Proposed Low

```{r}
proposed_lowincome_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_hh_income_low),color=NA)

proposed_lowincome_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Low Income Households",
                       breaks = size_brks <- seq(0, 2500, by = 500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("proposed_lowincome.pdf", 
  proposed_lowincome_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Low Income Households",
                       breaks = size_brks <- seq(0, 2500, by = 500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

####Existing Mid

```{r}
existing_midincome_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = hh_income_mid),color=NA)

existing_midincome_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Mid Income Households",
                       breaks = size_brks <- seq(0, 2000, by = 500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("existing_midincome.pdf", 
  existing_midincome_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Mid Income Households",
                       breaks = size_brks <- seq(0, 2000, by = 500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

####Proposed Mid

```{r}
proposed_midincome_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_hh_income_mid),color=NA)

proposed_midincome_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Mid Income Households",
                       breaks = size_brks <- seq(0, 6000, by = 1000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))
ggsave("proposed_midincome.pdf", 
  proposed_midincome_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Mid Income Households",
                       breaks = size_brks <- seq(0, 6000, by = 1000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

####Existing High

```{r}
existing_highincome_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = hh_income_high),color=NA)

existing_highincome_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "High Income Households",
                       breaks = size_brks <- seq(0, 3500, by = 1000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("existing_highincome.pdf", 
  existing_highincome_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "High Income Households",
                       breaks = size_brks <- seq(0, 3500, by = 1000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

####Proposed High

```{r}
proposed_highincome_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_hh_income_high),color=NA)

proposed_highincome_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed High Income Households",
                       breaks = size_brks <- seq(0, 12000, by = 2000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("propose_highincome.pdf", 
  proposed_highincome_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed High Income Households",
                       breaks = size_brks <- seq(0, 12000, by = 2000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

###Zero Car Households

####Existing Zero Car Households

```{r}
existing_nocar_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = nocar),color=NA)

existing_nocar_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Zero Car Households",
                       breaks = size_brks <- seq(0, 1200, by = 200),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("existing_nocar.pdf", 
  existing_nocar_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Zero Car Households",
                       breaks = size_brks <- seq(0, 1200, by = 200),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

####Proposed Zero Car Households

```{r}
proposed_nocar_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_nocar),color=NA)

proposed_nocar_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Zero Car Households",
                       breaks = size_brks <- seq(0, 2000, by = 500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("proposed_nocar.pdf", 
  proposed_nocar_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Zero Car Households",
                       breaks = size_brks <- seq(0, 2000, by = 500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

###Household Size

####Existing One Person Households

```{r}
existing_oneperson_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = oneperson_hh),color=NA)

existing_oneperson_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "One-Person Households",
                       breaks = size_brks <- seq(0, 3000, by = 500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("existing_oneperson.pdf", 
  existing_oneperson_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "One-Person Households",
                       breaks = size_brks <- seq(0, 3000, by = 500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

####Proposed One Person Households

```{r}
proposed_oneperson_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_oneperson_hh),color=NA)

proposed_oneperson_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed One-Person Households",
                       breaks = size_brks <- seq(0, 8000, by = 1200),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("proposed_oneperson.pdf",
  proposed_oneperson_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed One-Person Households",
                       breaks = size_brks <- seq(0, 8000, by = 1200),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

####Existing Two Person Households

```{r}
existing_twoperson_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = twoperson_hh),color=NA)

existing_twoperson_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Two-Person Households",
                       breaks = size_brks <- seq(0, 2500, by = 500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("existing_twoperson.pdf", 
  existing_twoperson_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Two-Person Households",
                       breaks = size_brks <- seq(0, 2500, by = 500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

####Proposed Two Person Households

```{r}
proposed_twoperson_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_twoperson_hh),color=NA)

proposed_twoperson_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Two-Person Households",
                       breaks = size_brks <- seq(0, 10000, by = 2000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("proposed_twoperson.pdf", 
  proposed_twoperson_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Two-Person Households",
                       breaks = size_brks <- seq(0, 10000, by = 2000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

###Employment

####Existing Basic Employees

```{r}
existing_basic_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = basic_emp),color=NA)

existing_basic_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Basic Employees",
                       breaks = size_brks <- seq(0, 15000, by = 3000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees"))

ggsave("existing_basic.pdf",
  existing_basic_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Basic Employees",
                       breaks = size_brks <- seq(0, 15000, by = 3000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees")))
```

####Proposed Basic Employees

```{r}
proposed_basic_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_basic),color=NA)

proposed_basic_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Basic Employees",
                       breaks = size_brks <- seq(0, 30000, by = 5000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees"))

ggsave("proposed_basic.pdf",
  proposed_basic_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Basic Employees",
                       breaks = size_brks <- seq(0, 30000, by = 5000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees")))
```

####Existing Service Employees

```{r}
existing_service_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = service_emp),color=NA)

existing_service_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Service Employees",
                       breaks = size_brks <- seq(0, 60000, by = 10000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees"))

ggsave("existing_service.pdf",
  existing_service_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Service Employees",
                       breaks = size_brks <- seq(0, 60000, by = 10000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees")))
```

####Proposed Service Employees

```{r}
proposed_service_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_service),color=NA)

proposed_service_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Service Employees",
                       breaks = size_brks <- seq(0, 90000, by = 15000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees"))

ggsave("proposed_service.pdf",
  proposed_service_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Service Employees",
                       breaks = size_brks <- seq(0, 90000, by = 15000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees")))
```

####Existing Retail Employees

```{r}
existing_retail_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = retail_emp),color=NA)

existing_retail_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Retail Employees",
                       breaks = size_brks <- seq(0, 5000, by = 1000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees"))

ggsave("existing_retail.pdf",
  existing_retail_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Retail Employees",
                       breaks = size_brks <- seq(0, 5000, by = 1000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees")))
```

####Proposed Retail Employees

```{r}
proposed_retail_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_retail),color=NA)

proposed_retail_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Retail Employees",
                       breaks = size_brks <- seq(0, 15000, by = 2500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees"))

ggsave("proposed_retail.pdf", 
  proposed_retail_map +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Retail Employees",
                       breaks = size_brks <- seq(0, 15000, by = 2500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees")))
```

###Summary Tables

We concluded by also determing the average, standard deviation, and median value of each variable in any given track. The tables demonstrate existing and proposed conditions.  

####Existing

```{r}
final_tracts %>%
  pivot_longer(cols=c(nocar,
                      population,
                      housing_units,
                      hh_income_low,
                      hh_income_mid,
                      hh_income_high,
                      oneperson_hh,
                      twoperson_hh,
                      basic_emp, 
                      retail_emp, 
                      service_emp),
               names_to = "Variable",
               values_to = "value") %>%
  group_by(Variable) %>%
  summarize(Average = mean(value, na.rm=TRUE),
            `Standard deviation` = sd(value, na.rm=TRUE),
            Median = median(value, median(value, na.rm=TRUE)))
```

####Proposed

```{r}
final_tracts %>%
  pivot_longer(cols=c(new_nocar,
                      new_total_tract_population,
                      new_total_housing_units,
                      new_hh_income_low,
                      new_hh_income_mid,
                      new_hh_income_high,
                      new_oneperson_hh,
                      new_twoperson_hh,
                      new_basic, 
                      new_retail, 
                      new_service),
               names_to = "Variable",
               values_to = "value") %>%
  group_by(Variable) %>%
  summarize(Average = mean(value, na.rm=TRUE),
            `Standard deviation` = sd(value, na.rm=TRUE),
            Median = median(value, median(value, na.rm=TRUE)))
```