---
title: "Comparing Existing and Proposed Conditions with Maps and Tables"
output: html_document
---
## Load Libraries

```{r}
library(ggplot2)
library(sf)
library(tidyverse)
library(tidycensus)
library(sp)
library(units)
library(here)
library(png)
library(ggthemes)
```

### Bringing in our Combined Census Tract Data

Here, we opened the csv file that we completed at the end of our second markdown file, TractZoneOverlay.Rmd. We also changed the values in the Proposed conditions variables to be numeric values.

```{r}
combinedtracts <- here("alternative",
                      "data",
                      "combined_cities",
                      "combined_tracts1.csv") %>%
  
  st_read() %>%
  mutate(new_basic = 
           as.numeric(new_basic),
         new_retail = 
           as.numeric(new_retail),
         new_service = 
           as.numeric(new_service),
         new_oneperson_hh = 
           as.numeric(new_oneperson_hh),
         new_twoperson_hh = 
           as.numeric(new_twoperson_hh),
         new_total_tract_population = 
           as.numeric(new_total_tract_population),
         new_total_housing_units = 
           as.numeric(new_total_housing_units),
         new_hh_income_low = 
           as.numeric(new_hh_income_low),
         new_hh_income_mid = 
           as.numeric(new_hh_income_mid),
         new_hh_income_high = 
           as.numeric(new_hh_income_high),
         new_nocar = 
           as.numeric(new_nocar))
         
  
```

### Selecting Variables to then Combine With Existing Conditions Data

To ensure that our new proposed condition variables would match with existing conditions data by tract, we created a new dataframe that contained the GEOID of our new proposed conditions and only the final counts of our new variables.

```{r}
final_tracts_sel <- combinedtracts %>%
  select(GEOID, 
         new_oneperson_hh,
         new_twoperson_hh,
         new_total_tract_population,
         new_total_housing_units,
         new_hh_income_low,
         new_hh_income_mid,
         new_hh_income_high,
         new_nocar,
         new_basic,
         new_retail,
         new_service)

write_csv(final_tracts_sel, here("alternative",
                                 "data",
                                 "proposed_zones.csv"))
```

### Combining Existing Conditions Tract Data with Proposed Conditions Tract Data

Last, we combined the Existing Conditions Tract Data from our first markdown file, ExistingConditions.Rmd, with the Proposed Conditions Data from our second markdown file, TractZoneOverlay.Rmd. They joined by GEOID.


```{r}
final_tracts <- left_join(zones, final_tracts_sel)
```




##Visualizing Existing and Proposed Conditions

Finally, we displayed comparison maps for each variable which compared the existing conditions of our tracts to the proposed conditions of our tracts. We did this using ggplot. We also saved each map as a pdf file for later addition to our report. 

###Population

####Existing Population

```{r}
existing_pop_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = population),color=NA)

existing_pop_map + 
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Tract Population",
                       breaks = size_brks <- seq(0, 40000, by = 10000),
                       limits = c(1, 40000), 
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "People"))

ggsave("population_tract.pdf", 
  existing_pop_map +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Tract Population",
                       breaks = size_brks <- seq(0, 40000, by = 10000),
                       limits = c(1, 40000),  
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "People")))
```
####Proposed Population

```{r}
proposed_pop_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_total_tract_population),color=NA)

proposed_pop_map + 
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Tract\nPopulation",
                       breaks = size_brks <- seq(0, 40000, by = 10000),
                       limits = c(1,40000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "People"))

ggsave("proposed_population.pdf", 
  proposed_pop_map + 
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Tract\nPopulation",
                       breaks = size_brks <- seq(0, 40000, by = 10000),
                       limits = c(1,40000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "People")))
```
###Households

####Existing Households
```{r}
existing_hh_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = housing_units),color=NA)

existing_hh_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Tract Housing Units",
                       breaks = size_brks <- seq(0, 15000, by = 2500),
                       limits = c(1, 16000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Units"))

ggsave("existing_hh.pdf", 
  existing_hh_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Tract Housing Units",
                       breaks = size_brks <- seq(0, 15000, by = 2500),
                       limits = c(1, 16000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Units")))
```

####Proposed Households

```{r}
proposed_hh_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_total_housing_units),color=NA)

proposed_hh_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Tract\nHousing Units",
                       breaks = size_brks <- seq(0, 15000, by = 2500),
                       limits = c(1, 16000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Units"))

ggsave("proposed_hh.pdf", 
  proposed_hh_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Tract\nHousing Units",
                       breaks = size_brks <- seq(0, 15000, by = 2500),
                       limits = c(1, 16000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Units")))
```

###Household Incomes

####Existing Low

```{r}
existing_lowincome_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = hh_income_low),color=NA)

existing_lowincome_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Low Income\nHouseholds",
                       breaks = size_brks <- seq(0, 2500, by = 500),
                       limits = c(1,2500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("existing_lowincome.pdf", 
  existing_lowincome_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Low Income\nHouseholds",
                       breaks = size_brks <- seq(0, 2500, by = 500),
                       limits = c(1,2500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

####Proposed Low

```{r}
proposed_lowincome_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_hh_income_low),color=NA)

proposed_lowincome_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Low Income\nHouseholds",
                       breaks = size_brks <- seq(0, 2500, by = 500),
                       limits = c(1,2500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("proposed_lowincome.pdf", 
  proposed_lowincome_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Low Income\nHouseholds",
                       breaks = size_brks <- seq(0, 2500, by = 500),
                       limits = c(1,2500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

####Existing Mid

```{r}
existing_midincome_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = hh_income_mid),color=NA)

existing_midincome_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Mid Income\nHouseholds",
                       breaks = size_brks <- seq(0, 6000, by = 1000),
                       limits = c(1,6000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("existing_midincome.pdf", 
  existing_midincome_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Mid Income\nHouseholds",
                       breaks = size_brks <- seq(0, 6000, by = 1000),
                       limits = c(1,6000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

####Proposed Mid

```{r}
proposed_midincome_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_hh_income_mid),color=NA)

proposed_midincome_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Mid Income\nHouseholds",
                       breaks = size_brks <- seq(0, 6000, by = 1000),
                       limits = c(1,6000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))
ggsave("proposed_midincome.pdf", 
  proposed_midincome_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Mid Income\nHouseholds",
                       breaks = size_brks <- seq(0, 6000, by = 1000),
                       limits = c(1,6000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

####Existing High

```{r}
existing_highincome_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = hh_income_high),color=NA)

existing_highincome_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "High Income\nHouseholds",
                       breaks = size_brks <- seq(0, 12000, by = 2000),
                       limits = c(1,12000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("existing_highincome.pdf", 
  existing_highincome_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "High Income\nHouseholds",
                       breaks = size_brks <- seq(0, 12000, by = 2000),
                       limits = c(1,12000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

####Proposed High

```{r}
proposed_highincome_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_hh_income_high),color=NA)

proposed_highincome_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed High Income\nHouseholds",
                       breaks = size_brks <- seq(0, 12000, by = 2000),
                       limits = c(1,12000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("propose_highincome.pdf", 
  proposed_highincome_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed High Income\nHouseholds",
                       breaks = size_brks <- seq(0, 12000, by = 2000),
                       limits = c(1,12000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

###Zero Car Households

####Existing Zero Car Households

```{r}
existing_nocar_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = nocar),color=NA)

existing_nocar_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Zero Car Households",
                       breaks = size_brks <- seq(0, 1800, by = 300),
                       limits = c(0,1800),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("existing_nocar.pdf", 
  existing_nocar_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Zero Car Households",
                       breaks = size_brks <- seq(0, 1800, by = 300),
                       limits = c(0,1800),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

####Proposed Zero Car Households

```{r}
proposed_nocar_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_nocar),color=NA)

proposed_nocar_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Zero\nCar Households",
                       breaks = size_brks <- seq(0, 1800, by = 300),
                       limits = c(0,1800),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("proposed_nocar.pdf", 
  proposed_nocar_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Zero\nCar Households",
                       breaks = size_brks <- seq(0, 1800, by = 300),
                       limits = c(0,1800),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

###Household Size

####Existing One Person Households

```{r}
existing_oneperson_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = oneperson_hh),color=NA)

existing_oneperson_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "One-Person Households",
                       breaks = size_brks <- seq(0, 7500, by = 1500),
                       limits = c(1,7500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("existing_oneperson.pdf", 
  existing_oneperson_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "One-Person Households",
                       breaks = size_brks <- seq(0, 7500, by = 1500),
                       limits = c(1,7500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

####Proposed One Person Households

```{r}
proposed_oneperson_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_oneperson_hh),color=NA)

proposed_oneperson_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed One-Person\nHouseholds",
                       breaks = size_brks <- seq(0, 7500, by = 1500),
                       limits = c(1,7500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("proposed_oneperson.pdf",
  proposed_oneperson_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed One-Person\nHouseholds",
                       breaks = size_brks <- seq(0, 7500, by = 1500),
                       limits = c(1,7500),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

####Existing Two Person Households

```{r}
existing_twoperson_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = twoperson_hh),color=NA)

existing_twoperson_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Two-Person Households",
                       breaks = size_brks <- seq(0, 10000, by = 2000),
                       limits = c(1,10000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("existing_twoperson.pdf", 
  existing_twoperson_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Two-Person Households",
                       breaks = size_brks <- seq(0, 10000, by = 2000),
                       limits = c(1,10000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

####Proposed Two Person Households

```{r}
proposed_twoperson_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_twoperson_hh),color=NA)

proposed_twoperson_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Two-Person\nHouseholds",
                       breaks = size_brks <- seq(0, 10000, by = 2000),
                       limits = c(1,10000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households"))

ggsave("proposed_twoperson.pdf", 
  proposed_twoperson_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Two-Person\nHouseholds",
                       breaks = size_brks <- seq(0, 10000, by = 2000),
                       limits = c(1,10000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Households")))
```

###Employment

####Existing Basic Employees

```{r}
existing_basic_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = basic_emp),color=NA)

existing_basic_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Basic Employees",
                       breaks = size_brks <- seq(0, 30000, by = 5000),
                       limits = c(1,31000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees"))

ggsave("existing_basic.pdf",
  existing_basic_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Basic Employees",
                       breaks = size_brks <- seq(0, 30000, by = 5000),
                       limits = c(1,31000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees")))
```

####Proposed Basic Employees

```{r}
proposed_basic_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_basic),color=NA)

proposed_basic_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Basic\nEmployees",
                       breaks = size_brks <- seq(0, 30000, by = 5000),
                       limits = c(1,31000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees"))

ggsave("proposed_basic.pdf",
  proposed_basic_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Basic\nEmployees",
                       breaks = size_brks <- seq(0, 30000, by = 5000),
                       limits = c(1,31000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees")))
```

####Existing Service Employees

```{r}
existing_service_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = service_emp),color=NA)

existing_service_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Service Employees",
                       breaks = size_brks <- seq(0, 85000, by = 15000),
                       limits = c(1,85000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees"))

ggsave("existing_service.pdf",
  existing_service_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Service Employees",
                       breaks = size_brks <- seq(0, 85000, by = 15000),
                       limits = c(1,85000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees")))
```

####Proposed Service Employees

```{r}
proposed_service_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_service),color=NA)

proposed_service_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Service\nEmployees",
                       breaks = size_brks <- seq(0, 85000, by = 15000),
                       limits = c(1,85000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees"))

ggsave("proposed_service.pdf",
  proposed_service_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Service\nEmployees",
                       breaks = size_brks <- seq(0, 85000, by = 15000),
                       limits = c(1,85000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees")))
```

####Existing Retail Employees

```{r}
existing_retail_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = retail_emp),color=NA)

existing_retail_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Retail Employees",
                       breaks = size_brks <- seq(0, 15000, by = 2500),
                       limits = c(1,15000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees"))

ggsave("existing_retail.pdf",
  existing_retail_map +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  theme_void() +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Retail Employees",
                       breaks = size_brks <- seq(0, 15000, by = 2500),
                       limits = c(1,15000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees")))
```

####Proposed Retail Employees

```{r}
proposed_retail_map <- ggplot(final_tracts)+
  geom_sf(aes(fill = new_retail),color=NA)

proposed_retail_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Retail\nEmployees",
                       breaks = size_brks <- seq(0, 15000, by = 2500),
                       limits = c(1,15000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees"))

ggsave("proposed_retail.pdf", 
  proposed_retail_map +
  theme_void() +
  theme(legend.title = element_text(size=10),
        legend.text=element_text(size=8)) +
  scale_fill_viridis_c(option = "D",
                       begin = 1, end = 0,
                       name = "Proposed Retail\nEmployees",
                       breaks = size_brks <- seq(0, 15000, by = 2500),
                       limits = c(1,15000),
                       labels = paste(formatC(size_brks,
                                              format="f",
                                              digits=0),
                                      "Employees")))
```

###Summary Tables

We concluded by also determing the average, standard deviation, and median value of each variable in any given track. The tables demonstrate existing and proposed conditions.  

####Existing

```{r}
final_tracts %>%
  pivot_longer(cols=c(nocar,
                      population,
                      housing_units,
                      hh_income_low,
                      hh_income_mid,
                      hh_income_high,
                      oneperson_hh,
                      twoperson_hh,
                      basic_emp, 
                      retail_emp, 
                      service_emp),
               names_to = "Variable",
               values_to = "value") %>%
  group_by(Variable) %>%
  summarize(Average = mean(value, na.rm=TRUE),
            `Standard deviation` = sd(value, na.rm=TRUE),
            Median = median(value, median(value, na.rm=TRUE)))
```

####Proposed

```{r}
final_tracts %>%
  pivot_longer(cols=c(new_nocar,
                      new_total_tract_population,
                      new_total_housing_units,
                      new_hh_income_low,
                      new_hh_income_mid,
                      new_hh_income_high,
                      new_oneperson_hh,
                      new_twoperson_hh,
                      new_basic, 
                      new_retail, 
                      new_service),
               names_to = "Variable",
               values_to = "value") %>%
  group_by(Variable) %>%
  summarize(Average = mean(value, na.rm=TRUE),
            `Standard deviation` = sd(value, na.rm=TRUE),
            Median = median(value, median(value, na.rm=TRUE)))
```